---
title: "pca, regression, and pls"
output: html_notebook
---

call the libraries

```{r message=FALSE}
library(tidymodels)
library(data.table)
library(magrittr)
library(ggplot2)
library(GGally)
```

Let's get a dataset we can work with

```{r}
data(biomass)
dt <- data.table(biomass)
rm(biomass)
```


let's get a look at it

```{r}
dt
str(dt)
```

are these guys related to each other?  what should we try and predict?

```{r}
dt[, -c('sample', 'dataset')] %>% ggpairs()
```

carbon and HHV are quite closely related.  Let's leave that out for this analysis.

what does a standard, run-of-the-mill regression look like?

```{r}
native_regression <- lm(HHV~., data=dt[, -c('sample', 'dataset', 'carbon')]) 
native_regression %>% summary()
```
```{r}
native_pred <- predict(native_regression, newdata=dt[, -c('sample', 'dataset', 'carbon')])
data.table(truth = dt$HHV, pred = native_pred) %>% ggplot(aes(x=truth, y=pred)) + geom_point()
```
```


now, what would pca look like

```{r}
pca_model <- prcomp(dt[, -c('sample', 'dataset', 'carbon', 'HHV')], center = TRUE, scale. = TRUE)
summary(pca_model)
```

```{r}
pca_dt <- predict(pca_model, newdata=dt[, -c('sample', 'dataset', 'carbon', 'HHV')]) %>% data.table()
pca_dt
```
what does a regression look like with those first couple or three pc's?

```{r}
pca_dt[, HHV:=dt$HHV]
pca_regression <- lm(HHV~., data=pca_dt[, -c('PC3')])
pca_regression %>% summary()
```

weird.  I don't know why it's #3 that's the easy one to drop

```{r}
pca_pred <- predict(pca_regression, newdata=pca_dt)
data.table(truth = dt$HHV, pred = pca_pred) %>% ggplot(aes(x=truth, y=pred)) + geom_point()
```


now, let's try PLS, which could stand for "projection onto latent structures"

```{r}

```

